---
title: "Process Notebook"
author: "Connie Xu, Arielle Herman, Zhirui (Ariel) Luo, Grace Li"
date: "3/18/2021"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
# install.packages (basic)
suppressMessages(library(dplyr))
suppressMessages(library(tidyverse))

# install.packages (reading)
suppressMessages(library(XML))
suppressMessages(library(RCurl))
suppressMessages(library(readr))
suppressMessages(library("readxl"))

# install.packages (themes)
suppressMessages(library(ggthemes))
suppressMessages(library(ggrepel))
suppressMessages(library(RColorBrewer))
suppressMessages(library(viridis))
suppressMessages(library(hrbrthemes))
suppressMessages(library(plotly))


# install.packages (maps)
suppressMessages(library(RgoogleMaps))
suppressMessages(library(ggmap))
suppressMessages(install.packages("maps"))
suppressMessages(install.packages("tmap")) # install the CRAN version
suppressMessages(library(tmap))
suppressMessages(install.packages('rgeos'))

suppressMessages(library(devtools))
# Let's install the development version from Github. Run
devtools::install_github("rstudio/leaflet")
```
## Introduction 
As discussed in other sections of our repository, this project sought to explore a factors related to socioeconomic conditions, student debt, inequities, and policy. This was a realy broad topic so we did a decent amount of exploration of the data prior to deciding on a more specific topic. 

## The Data 
We first worked with the 2019 College Scorecard to look at some trends that we thought would be interesting. The following is the dataset and some basic columns added to it for our analysis work. 
```{r csv load, eval=TRUE, echo=FALSE, message = TRUE, warnings=FALSE}
# Set the WD as Group_G_HigherEd 
setwd("~/Dropbox (Business)/Spring 2021/QMSS 5063 - Data Visualization /Group_G_HigherEd")

# While our initial College Scorecard only included 2019 initially, I was able to run a simple python script to concatenate and select a small number of relevant columns for our visualization and analysis. Thus, as our first visualizations only unclude 2019, the code below is for re-filtering the concatenated 2010-2019 data back into simply 2019. 

sc_time <- read_csv('src/2010_2019_student_debt.csv') 
sc <- sc_time %>% filter(Year_Ending == 2019)

# Renaming Latitude and Longitude Columns for later use 
sc <- sc %>% dplyr::rename(lat = LATITUDE) %>% dplyr::rename(long = LONGITUDE) %>% dplyr::rename(state = STABBR)

#recoding for admissions rates (ADM_RATE) to note selectivity level of schools
sc <- sc %>% dplyr::mutate(uni_rank = ifelse(ADM_RATE<0.05,'elite',ifelse(ADM_RATE<0.2,'highly selective', ifelse(ADM_RATE<0.3, 'selective',ifelse(ADM_RATE<0.7,'less selective', 'not selective')))))
colnames(sc)
```


### Selected Columns (Described in College Scorecard)  

| variable                 | description                                               |
| :----------------------- | --------------------------------------------------------- |
| UNITID                   | IPEDS Unit ID                                             |   
| INSTNM                   | Institution Name                                          |
| CITY                     | description                                               |
| state                    | state                                                     |
| ZIP                      | Zip Code                                                  |
| DEBT_MDN                 | The median original amount of the loan principal upon entering repayment|
| UGDS                     | Number 
| ADM_RATE                 | Admission Rate                                            | 
| CONTROL                  | Private / Public Nonprofit (1 and 2) vs. For-Profit (3) Schools|
| MN_EARN_WNE_P10          | Min Earning after 10 Years (not used due to incomplete data) |
| MD_EARN_WNE_P10          | Med Earning after 10 Years (not used due to incomplete data) |
| UGDS_WHITE; UGDS_BLACK; UGDS_HISP; UGD.S_ASIAN | Proportion of Undergrads (White, Black, Hispanic, Asian)| 
| UGDS                     | Number of Undergrads                                      | 
| School                   | Same as INSTNM (disregard - used to filter out community colleges) | 
| ADM_RATE_ALL_1           | Admission Rate by OPE ID (not used in the end)            | 
| Year, Year_Starting, Year_Ending | Year Starting (twice due to how the data was read in) and Year Ending period covered by the scorecard | 

### Additional Filtering / Column Mutation

```{r}
# Selectivity Categories were made guided by as-yet unpublished work from Conwell et al. (2022 exp.) [1] as well as by articles from the US News and Rankings 
sc <- sc %>%
  dplyr::mutate(uni_rank = case_when(
    ADM_RATE < 0.05 ~ 'elite',
    ADM_RATE < 0.2 ~ 'highly selective',
    ADM_RATE < 0.3 ~ 'more selective',
    ADM_RATE < 0.5 ~ 'selective',
    ADM_RATE < 0.7 ~ 'less selective',
    TRUE ~ 'not selective')) %>% mutate(uni_rank = factor(uni_rank, levels=c('not selective', 'less selective', 'selective', 'more selective', 'highly selective', 'elite')))
```

## Preliminary (Very Rough) Drafts - Searching for Trends 

As discussed in the introductory section, we had far too many subtopics within our dataset and we wanted to better explore a specific sub-topic. Thus, we began with trying to identify patterns within our dataset and other related datasets. 

We started by testing whether or not we would be able to generate a graph of where universities in the US are located. This was mostly to validate that the lat and lon columns were working properly and were not going to cause issues if we were to plot these institutions on a map. 

```{r initial map matching, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
#note 
us <- map_data("state")
us_states <- as_data_frame(us) %>% dplyr::rename(state=region)
us_states

mapping=aes(x = lat, y = long)
ggplot(us,  aes(x = long, y = lat, group=group)) +
  geom_polygon(aes(), color='white') + 
  theme_ipsum() + theme(panel.grid = element_blank())+ geom_point(data = df, aes(x = long, y = lat, group=UNITID),color='gold',alpha=0.5) 

```
We then started trying to look at some trends that could be compiled into our visualization project. Some themes we were interested in were *race*, *socioeconomic status*, *student debt*, and *economic outcomes*.  
```{r competitiveness_student_debt, echo=TRUE,eval=TRUE, message=FALSE,warning=FALSE}
sc_sub <- sc %>% subset(DEBT_MDN != 'PrivacySuppressed') %>% drop_na(uni_rank) %>% drop_na(DEBT_MDN)
sc_sub$uni_rank_cat = factor(sc_sub$uni_rank, levels=c('elite','highly selective','selective','less selective','not selective'))

ggplot(sc_sub, aes(x=uni_rank_cat,y=DEBT_MDN)) +
  geom_point(color = '#87495F',alpha=.12) +
  theme_ipsum(base_size = 12, axis_title_size = 14) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
      ggtitle("Student Debt and School Tiers") +    
  labs(fill = "Selectivity", y="Student Debt", x = "School Competitiveness")

```
  
This was not a very good visual but I wanted to see whether student debt varied relative to the 'eliteness' and the 'selectiveness' of the school. As can be somewhat seen above, elite schools and highly selective schools appear to potentially exhibit higher student debt on average.  

The next thing we did was look the percentage of Black students vs. the Median Debt. 

```{r debt_race, echo=TRUE,eval=TRUE, message=FALSE,warning=FALSE}
#dfsub %>% group_by(uni_rank) mutate(density = n()) 

m <- ggplot(dfsub, aes(x=DEBT_MDN,y=UGDS_BLACK, color = uni_rank_cat)) +
  geom_point(aes(color = uni_rank_cat), alpha=.4) +
  scale_color_manual(values=c("navy","lightblue","aquamarine3","orange","gold"))+
  geom_smooth(method = lm, se = FALSE)+
  theme_ipsum() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
      ggtitle("Student Debt and Racial Composition (Black Students)") +    
  labs(y="% Black Students", x = "Median Student Debt ($)")
ggplotly(m)
m
```
  
Again this is poorly presented (and as I recognized later, the major issue with all of the grey at the axis was due to an overcrowded axis label such that we would be unable to see what the x and y values included). This also did not include a very clear trend - as such, I left 'race and student debt' alone for the time being as we were attempting to work on scaling down to more specific topics. 

---  
[1] - Conwell, Jordan A. and Natasha Quadlin. Forthcoming. “Race, Gender,Higher Education, and Socioeconomic Attainment: Evidence from Baby Boomers at Mid-Life.” Social Forces. DOI: 10.1093/sf/soab010.

