---
title: "shiny_attempt1"
author: "Arielle Herman"
date: "4/18/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(shinydashboard)
library(tidyverse)
library(ggridges)
```


```{r}
ui <- fluidPage(title = "Random generator",
  tabsetPanel(              
    tabPanel(title = "Final",
      plotOutput("final")),
    navbarMenu(title = "other versions",
    tabPanel(title = "Uniform data",
      plotOutput("violin")),
    tabPanel(title = "Chi Squared data",
      plotOutput("ggridges"))
     )
  )
)

```

```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Group G: Student Debt"),
  dashboardSidebar(
    sidebarMenu(
      menuItem("About", tabName = "about"),
      menuItem("Student Debt", tabName = "debt"),
      menuItem("What do people think?", tabName = "tweets")
    )
  ),
  dashboardBody(
    tabItems(
      tabItem(tabName = "about", class = "active",
              fluidPage()),
      tabItem(tabName = "debt", class = "active",
        fluidPage(
          wellPanel(plotOutput("final")),
          wellPanel(plotOutput("violin")),
          wellPanel(plotOutput("ggridges"))
        )
      ),
      tabItem(tabName = "tweets", class = "active",
          fluidPage()
      )
    )
  )
)


#server <- function(input, output) { }

shinyApp(ui, server)
```

```{r}
ui <- fluidPage(title = "practice",
                navlistPanel(
                  "About",
                  tabPanel("Project Motivation"),
                  tabPanel("Data"),
                  tabPanel("Team Members"),
                  "Data Visuals",
                  tabPanel("Repayment Rates",
                           plotOutput("final"),
                           plotOutput("violin"),
                           plotOutput("ggridges")),
                  tabPanel("Student Debt"),
                  tabPanel("Loan Forgiveness Tweets"),
                  widths = c(5,10)
  ))

server <- function(input, output) {
  library(tidyverse)
  output$final <- renderPlot({

      sc_repay %>%
      filter(years_since_entering_repay == 1, repay_rate > 0) %>%
      
      # plot
      ggplot(aes(x = repay_rate, y = fct_rev(factor(cohort_year)), fill = stat(x))) +
      scale_fill_distiller(name = "Repay Rate", palette = "greens") +
      ggridges::geom_density_ridges_gradient(scale = 2, alpha = 0.50) +
      
      # labels
      ggtitle("Proportion of University Fiscal Cohorts Whose Loan\nBalances Declined in Year One of Repayment\n") +
      ylab("Fiscal Cohort Year\n") + xlab(NULL) +
      scale_x_continuous(labels = scales::dollar_format(suffix = "%", prefix = "")) +
      scale_x_continuous(labels = scales::percent_format()) +
      ourtheme + theme(legend.position = "none")
    
  })
  
  output$violin <- renderPlot({
    
    sc_repay %>%
      pivot_longer(cols = c("repay_rate_lowincome", "repay_rate_midincome", "repay_rate_highincome"),
               names_to = "class", values_to = "rates") %>%
      filter(rates > 0) %>%
      mutate(class = factor(class, levels =
                          c("repay_rate_highincome", "repay_rate_midincome", "repay_rate_lowincome"))) %>%
      
      ggplot(aes(x = year, y = rates, group = cohort_year, fill = class)) +
      geom_violin(color = "lightgray") + facet_grid(. ~ class) +
      ggthemes::theme_tufte() + # add base theme first and then modify
      theme(panel.spacing = unit(1, "cm"),
            panel.border = element_rect(colour = "lightgray", fill=NA, size=1),
            legend.position = "bottom",
            plot.title = element_text(hjust=0.5),
            strip.text.x = element_blank(), axis.ticks = element_blank()) +
      
      ## labels
      ggtitle("Fraction Of Students In Successive Fiscal Cohorts\nWhose Loan Balances Have Declined") +
      xlab(NULL) + ylab(NULL) +
      scale_fill_viridis_d(name = "Income Range:", labels = c("High Income", "Middle Income", "Low Income")) +
      ## percents
      scale_y_continuous(labels = scales::dollar_format(suffix = "%", prefix = "")) +
      scale_y_continuous(labels = scales::percent_format())
    
  })
  
  output$ggridges <- renderPlot({
    
    sc_repay %>% mutate(opeid = as.integer(opeid)) %>%
      left_join(selectivity, by = c("opeid" = "OPEID")) %>%
      filter(years_since_entering_repay == 1) %>%
      select(repay_rate, uni_rank, cohort_year, year, opeid) %>%
      mutate(admission_half = ifelse(uni_rank %in% c("elite", "highly selective", "more selective"), 1, 0)) %>%
      na.omit() %>%
      
      filter(repay_rate > 0, uni_rank != "not selective") %>%
  
      # plot
      ggplot(aes(x = repay_rate, y = fct_rev(factor(cohort_year)), fill = uni_rank)) +
      ggridges::geom_density_ridges(scale = 2, color = "darkgray", alpha = 0.50) +
      
      # labels
      ggtitle("Proportion of University Fiscal Cohorts\nWhose Loan Balances Declined") +
      ylab("Fiscal Cohort\n") + xlab(NULL) +
      scale_fill_viridis_d(name = "Selectivity:", labels = c("Less", "Selective", "More", "Highly"), direction = 1) +
      scale_x_continuous(labels = scales::dollar_format(suffix = "%", prefix = "")) +
      scale_x_continuous(labels = scales::percent_format()) +
      ourtheme
  })
}

shinyApp(server = server, ui = ui)

```
This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



